---
author: ""
date: ""
title: ""
output:
  html_document:
    toc: false
---

```{r setup, echo = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  # cache = TRUE,
  # include = FALSE,
  # results = "markdown",
  # results = "hide",
  fig.align = "center",
  # fig.show = "asis",
  fig.width = 6,
  fig.height = 6,
  # out.width = 6,
  # out.height = 6,
  warning = FALSE,
  message = FALSE
)

```


```{r load}
source(file.path("R", "00-config.R"))
load(params$path_analysis_image)
```

```
<script>
  $(".toggle").click(function() {
    $(this).toggleClass("open");
  });
</script>
```

<div class="toggle"><button>Show</button>
</div>


__NOTE:__
<i>
This is part of a series of write-ups discussing my findings of
[Texas high school academic University Interscholastic Scholarship (UIL) competitions](http://www.uiltexas.org/academics).
See the first write-up [here](/posts/2018-05-13-analysis-texas-high-schools-academics-r).

To keep this and the other write-ups concise and to focus reader attention on the content,
I have decided not to show the underlying code
(especially that which is used to create the visuals). Nonetheless, the full code
can be viewed on [my GitHub account](https://github.com/tonyelhabr). In the future, I may write some
some kind of addendum to demonstrate some of the programming implementation that I
think is notable in some way.
</i>

## Introduction


Ever since I finished high school in 2012, I've wanted to look back and evaluate my
performance in the academic [University Interscholastic League](http://www.uiltexas.org/) (UIL)
competitions that I competed in with historical results.
(To provide some background, most public high schools are registered in 
the [University Interscholastic League](http://www.uiltexas.org/) (UIL), which 
"exists to provide educational extracurricular academic, athletic, and music contests".
For those familiar with the [National Collegiate Athletic Association](https://www.ncaa.com/) (NCAA),
the UIL serves an analogous role for Texas high school extracurricular activities.)

Aside from my own self-interest in the historical results in these competitions,
I think that this analysis can provide some insight into which schools (and individual students)
are really the most "elite". State-wide and nation-wide standardized test scores
certainly are the most common measure of school and student academic strength,
[but they may not truly be the best]().

### About the Data [^source]

To make some sense of the my analysis, the reader should be aware of the following details about the data.

+ The competition data was scraped from https://www.hpscience.net/ for all years from 2008 through 2017.
The data is not listed in an extremely user-friendly format (in my opinion). Consequently, the "cleaned"
data is imperfect in some ways.

+ The UIL categorizes schools into one of six "conferences". The conference labels
range from `1A`, `2A`, ..., `6A`, where the increasing leading digit (i.e. `1`, `2`, etc.)
generally corresponds to increasing school size.

+ Schools only compete against other schools in their conference.

+ The UIL defines 3 levels of competition (in order of "difficulty"): _District_, _Region_, and _State_.
These are listed in order of "difficulty". That is, 
Winning a District competitions, results in a Region competition appearance,
and, subsequently, winning a Region competiton results in a State competition appearance.
(Keep in mind that schools still only compete against other schools in their same
conference, even as they advance.)

+ The UIL defines _32 total Districts_ in Texas, which are aggregated into _4 Regions_.
(The source of the geo-spatial data is
https://nces.ed.gov/opengis/rest/services/K12_School_Locations/EDGE_GEOCODE_PUBLICSCH_1516/MapServer.)


```{r viz_map_bycomplvl_grid}
gridExtra::grid.arrange(viz_map_bycomplvl_grid)
```

+ For schools, winning is a "winner-take-all" matter: only the school with the most combined points
among its individual competitors advances. On ther other hand, an individual
may advance even if his school does not win if he places among the top _n_.
The value of _n_ is dependent on the competition type. [^uil_rules]

+ There are 5 different academic competitions "types":
_Calculator Applications_, _Computer Science_, _Mathematics_, _Number Sense_, and _Science_. [^all_competitions]


[^uil_rules]:
See the [UIL rules]() for more details.

[^all_competitions]:
[There are many more UIL competition types](http://www.uiltexas.org/files/academics/academics-brief-overview.pdf)
than those analyzed here
(including competitons for theatre, band, etc.), but these are the ones for academics.



