---
author: ""
date: ""
title: ""
output:
  html_document:
    toc: false
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  # cache = TRUE,
  # include = FALSE,
  # results = "markdown",
  # results = "hide",
  fig.align = "center",
  # fig.show = "asis",
  fig.width = 6,
  fig.height = 6,
  # out.width = 6,
  # out.height = 6,
  warning = FALSE,
  message = FALSE
)

```

```{r funcs_general, include = FALSE}

```


```{r load, include = FALSE}
source(file.path("R", "config.R"))
# load(config$path_analysis_image)

schools_nces_join <-
  config$path_schools_uil_nces_join %>%
  teproj::import_path_cleanly()
```

<style>
.toggle {
  height: 1.85em;
  overflow-y: hidden;
}
.toggle.open {
  height: auto;
}
</style>
<script>
  $(".toggle").click(function() {
    $(this).toggleClass("open");
  });
</script>

<!--<div class="toggle"><button>Show</button>-->

__NOTE:__

This is part of a series of write-ups discussing my findings of
[Texas high school academic University Interscholastic Scholarship (UIL) competitions](http://www.uiltexas.org/academics).
See the first write-up [here](/posts/2018-05-13-analysis-texas-high-schools_uil-academics-r).

To keep this and the other write-ups concise and to focus reader attention on the content,
I have decided not to show the underlying code
(especially that which is used to create the visuals). Nonetheless, the full code
can be viewed on [my GitHub account](https://github.com/tonyelhabr). In the future, I may write some
some kind of addendum to demonstrate some of the programming implementation that I
think is notable in some way.

<!--</div>-->

## Introduction

Ever since I finished high school in 2012, I've wanted to look back and evaluate my
performance in the academic [University Interscholastic League](http://www.uiltexas.org/) (UIL)
competitions that I competed in with historical results.
(To provide some background, most public high schools_uil are registered in 
the [University Interscholastic League](http://www.uiltexas.org/) (UIL), which 
"exists to provide educational extracurricular academic, athletic, and music contests".
For those familiar with the [National Collegiate Athletic Association](https://www.ncaa.com/) (NCAA),
the UIL serves an analogous role for Texas high school extracurricular activities.)

Aside from my own self-interest in the historical results in these competitions,
I think that this analysis can provide some insight into which schools_uil (and individual students)
are really the most "elite". State-wide and nation-wide standardized test scores
certainly are the most common measure of school and student academic strength,
[but they may not truly be the best]().

### About the Data

To make some sense of the my analysis, the reader should be aware of the following details about the data.

+ The competition data was scraped from https://www.hpscience.net/ for all years from 2008 through 2017.
The data is not listed in an extremely user-friendly format (in my opinion). Consequently, the "cleaned"
data is imperfect in some ways.

+ The UIL categorizes schools_uil into one of six "conferences". The conference labels
range from 1A, 2A, ..., 6A, where the increasing leading digit (i.e. 1, 2, etc.)
generally corresponds to increasing school size.

+ Schools only compete against other schools_uil in their conference.

+ The UIL defines 3 levels of competition (in order of "difficulty"): _District_, _Region_, and _State_.
These are listed in order of "difficulty". That is, 
Winning a District competitions, results in a Region competition appearance,
and, subsequently, winning a Region competiton results in a State competition appearance.
(Keep in mind that schools_uil still only compete against other schools_uil in their same
conference, even as they advance.)

+ The UIL defines _32 total Districts_ in Texas, which are aggregated into _4 Regions_.
(The source of the geo-spatial data is
https://nces.ed.gov/opengis/rest/services/K12_School_Locations/EDGE_GEOCODE_PUBLICSCH_1516/MapServer.)


```{r viz_map_bycomplvl_grid}
visualize_map_bycomplvl <-
  function(data = NULL) {
    teplot::create_map_base_tx() +
      geom_polygon(
        data =
          data %>%
          count(county, complvl_num, complvl, sort = TRUE) %>%
          mutate_at(vars(complvl_num), funs(factor)) %>%
          inner_join(
            teplot::get_map_data_county_tx(),
            by = c("county" = "subregion")
          ),
        aes(group = county, fill = complvl_num),
      ) +
      viridis::scale_fill_viridis(# guide = guide_legend(title = complvls, nrow = 4, byrow = TRUE)
        discrete = T,
        option = "D") +
      theme(legend.position = "bottom")
  }

lab_title_map <-
  "Locations of Schools, by Competition Level"

lab_caption_map <-
  str_wrap(
    paste0(
      "Note that some counties may be incorrectly categorized ",
      "due to imperfect data extraction and cleaning. \n\n",
      "Data Source:\n",
      "The geo-spatial data was downloaded from ",
      "https://nces.ed.gov/opengis/rest/services/K12_School_Locations",
      "/EDGE_GEOCODE_PUBLICSCH_1516/MapServer ",
      "and combined with the school information (including districts and regions) ",
      "accompanying the UIL scores ",
      "scraped from https://www.hpscience.net/ via \"fuzzy-joining\"."
      ),
    60)

viz_map_byregion <-
  schools_nces_join %>%
  filter(complvl == "Region") %>%
  visualize_map_bycomplvl() +
  guides(fill = guide_legend(title = "Region", nrow = 1, byrow = FALSE)) +
  theme(plot.caption = element_text(hjust = 0)) +
  labs(title = lab_title_map, caption = lab_caption_map)
viz_map_byregion

viz_map_bydistrict <-
  schools_nces_join %>%
  filter(complvl == "District") %>%
  visualize_map_bycomplvl() +
  guides(fill = guide_legend(title = "District", nrow = 4, byrow = TRUE))
viz_map_bydistrict

viz_map_bycomplvl_grid <-
  gridExtra::arrangeGrob(
    viz_map_byregion,
    viz_map_bydistrict,
    nrow = 1
  )

# gridExtra::grid.arrange(viz_map_bycomplvl_grid)

teproj::export_ext_png(
  # viz_map_bycomplvl_grid,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 8,
  width = 10,
  viz_map_bycomplvl_grid
)
```

+ For schools_uil, winning is a "winner-take-all" matter: only the school with the most combined points
among its individual competitors advances. On ther other hand, an individual
may advance even if his school does not win if he places among the top _n_.
The value of _n_ is dependent on the competition type. [^uil_rules]

+ There are 5 different academic competitions "types":
_Calculator Applications_, _Computer Science_, _Mathematics_, _Number Sense_, and _Science_. [^all_competitions]


[^uil_rules]:
See the [UIL rules]() for more details.

[^all_competitions]:
[There are many more UIL competition types](http://www.uiltexas.org/files/academics/academics-brief-overview.pdf)
than those analyzed here
(including competitons for theatre, band, etc.), but these are the ones for academics.

## What's Next

In this series, I plan to investigate the following topics:

+ The relationships among the competition years, types, levels

+ Individual performance (including myself)

+ School performance (including my high school)

+ Miscellaneous questions, such as sibling performance

Performance evaluation relies primarily upon
test scores (from which other useful metrics may be derived, such as advancement).




