---
author: ""
date: ""
title: ""
output:
  html_document:
    toc: false
---

```{r setup, echo = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  # cache = TRUE,
  # include = FALSE,
  # results = "markdown",
  # results = "hide",
  fig.align = "center",
  # fig.show = "asis",
  fig.width = 10,
  fig.height = 10,
  # out.width = 6,
  # out.height = 6,
  warning = FALSE,
  message = FALSE
)

```

```{r load}
source(file.path("R", "00-config.R"))
load(params$path_analysis_image)
```

## Individual Participation

To begin, et's take a closer look at individual competitors. Which individuals
have competed the most?

```{r html_person_n_top}
# persons_n %>%
#   DT::datatable(
#     filter = "bottom"
#   )
html_person_n_top
```

Although the names here may not provide much insight, the counts provide some context
regarding the limits of individual participation.

Given that individual participation may not be indicative of anything directly,
it may be a better idea to break it down by conference. What is the count
of unique individuals competing in each conference?


```{r html_person_n_top_byconf}
html_person_n_top_byconf
```

It seems that there are not as many individual competitors in the 6A conference (`conf 6`).
This seems to be true even when stratifying the individuals into groups of ranks (by
volume of individual participation).

```{r viz_persons_n_top_byconf}
viz_persons_n_top_byconf
```

I would hypothesize that the noticeable lack of individual competitors in conference 6A--
which is the conference with largest high schools (according to student body size)--
can be attributed to "allocation of talent" by these schools.
In other words, these schools may be more likely to designate their individual
competitors to compete in only specific competitions.
By comparison, schools in the other conferences may be more willing to let individual
students to compete in as many competition types as they desire, even if they have
not prepared for them whatsoever.

Along these lines, note that there is a limit on the number of entrants
per school in a given competition. This means that smaller schools are much less likely to ever
have enough people to reach the per-school limit, so individuals at these schools
are welcome to compete in as many competitions as they would like.

Such a phenomenon
might be made evident by lower scores (in aggregate) for conferences where participation
by distinct individuals is greater. Nonetheless, this was not what 



