---
author: ""
date: ""
title: ""
output:
  html_document:
    toc: false
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  cache = TRUE,
  # include = FALSE,
  # results = "markdown",
  # results = "hide",
  fig.align = "center",
  # fig.show = "asis",
  fig.width = 7,
  fig.height = 7,
  # out.width = 7,
  # out.height = 7,
  warning = FALSE,
  message = FALSE
)

```

```{r load include = FALSE}
source(file.path("R", "config.R"))
schools_uil <-
  config$path_schools_uil %>%
  teproj::import_path_cleanly()

# persons_uil <-
#   config$path_persons_uil %>%
#   teproj::import_path_cleanly()

```

Having investigated individuals elsewhere, let's now take a look at the schools.

__NOTE:__
<i>
Although began the examinations of competitions and individuals by looking at
volume of participation (to provide context), 
I'll skip an analogous discussion here
because the participation of schools is shown indirectly through
those analyses.)
</i>

## School Scores

Let's begin by looking at some of the same metrics shown for invidual students,
but aggregated across all students for each school. In order to give the reader
some insight into school performance, I'll rank and show schools by a singular metric of performance
To be consistent,
I'll use the same
metric used for ranking the invdividuals--summed percentile rank of scores (`prnk_sum`).

__NOTE:__
<i>
For the same reason stated before for showing my own scores among the individuals,
I'll include the numbers for my high school ("CLEMENS") in applicable contexts.
</i>

```{r schools_stats, include = FALSE}
schools_stats <-
  schools_uil %>%
  add_schools_stats_cols_by_at(
    cols_grp = c("school")
  )
schools_stats %>% tetidy::count_arrange(n_state)
```

```{r html_schools_stats_filt, echo = FALSE, include = TRUE}
html_schools_stats_filt <-
  schools_stats %>%
  mutate_if(is.numeric, funs(round(., 2))) %>%
  create_kable_filt_at(
    config = config,
    col_rgx = "school"
  )
html_schools_stats_filt
```

Admittedly, there's not a lot of insight to extract from this summary regarding
individual schools. Nonetheless, it provides some useful context regarding
the magnitude of performance metric values aggregated at the school level.

To begin creating some insight into this list of top-performing schools,
let's break down school performance by year. Also,
let's combine the performance metric values
with coordinate data to visualize where the best schools are.

```{r schools_stats_byyear, include = FALSE}
schools_stats_byyear <-
  schools_uil %>%
  add_schools_stats_cols_by_at(
    cols_grp = c("school", "year")
  )
schools_stats_byyear
```

```{r html_schools_stats_byyear, echo = FALSE, include = TRUE, eval = FALSE}
html_schools_stats_byyear <-
  schools_stats_byyear %>%
  mutate_if(is.numeric, funs(round(., 2))) %>%
  create_kable_filt_at(
    config = config,
    col_rgx = "school"
  )
```


```{r visualize_map_tier3_at, include = FALSE}
add_tier3_lab_cols_at <-
  function(data = NULL,
           col = "rnk",
           col_out_chr = "rnk_tier_lab",
           tier_base = 3,
           n_tiers = 3,
           factor = TRUE,
           add_col_numeric = TRUE,
           col_out_num = "rnk_tier") {
    tier_max <- tier_base * 10 ^ (n_tiers - 1L - 1L)
    if (tier_max > nrow(data)) {
      warning("Too many tiers specified given rows in `data`.", call. = FALSE)
      return(data)
    }
    tiers <-
      tier_base * (10 ^ seq(1L - 1L, n_tiers - 1L - 1L, by = 1L))
    tier_lab_prefixes <-
      c(rep("Top ", n_tiers - 1L), "Outside top ")
    tier_labs <- paste0(tier_lab_prefixes, c(tiers, rev(tiers)[1]))

    col <- sym(col)
    col_out_chr <- sym(col_out_chr)
    ret <-
      data %>%
      mutate(
        !!col_out_chr :=
          case_when(
            !!col <= tiers[1] ~ tier_labs[1],
            !!col <= tiers[2] ~ tier_labs[2],
            TRUE ~ tier_labs[3]
          )
      )

    # NOTE: Factor first, then unfactor later if necessary.
    ret <-
      ret %>%
      mutate_at(vars(!!col_out_chr), funs(factor(., levels = rev(tier_labs))))

    if (add_col_numeric) {
      col_out_num <- sym(col_out_num)
      ret <-
        ret %>%
        mutate(!!col_out_num := as.integer(!!col_out_chr))
    }

    if (!factor) {
      ret <-
        ret %>%
        mutate_at(vars(!!col_out), funs(as.character(.)))
    }
    ret
  }

visualize_map_tier3_at <-
  function(data = NULL,
           col = "rnk",
           tier_base = 3,
           n_tier = 3,
           add_labels = TRUE,
           col_label = "school") {
    data_map <-
      # schools_stats %>%
      # inner_join(schools_nces_join %>% distinct(school, county, lat, lon), by = c("school")) %>%
      data %>%
      tetidy::rank_arrange_at(col) %>%
      add_tier3_lab_cols_at(tier_base = tier_base, n_tier = n_tier)

    rnk_tier_labs <- levels(data_map$rnk_tier_lab)

    data_tier1 <-
      data_map %>%
      filter(rnk_tier_lab == rev(rnk_tier_labs)[1])
    data_tier2 <-
      data_map %>%
      filter(rnk_tier_lab == rev(rnk_tier_labs)[2])
    data_tier3 <-
      data_map %>%
      filter(rnk_tier_lab == rev(rnk_tier_labs)[3])
    ret <-
      teplot::create_map_base_tx() +
      # geom_point(
      #   data = data_map,
      #   aes(
      #     x = lon,
      #     y = lat,
      #     color = rnk_tier_lab,
      #     size = rnk_tier,
      #     alpha = rnk_tier,
      #     group = ""
      #   )
      # ) +
    geom_point(
      data = data_tier3,
      aes(
        x = lon,
        y = lat,
        color = rnk_tier_lab,
        group = ""
      ),
      size = 1,
      alpha = 0.5
    ) +
      geom_point(
        data = data_tier2,
        aes(
          x = lon,
          y = lat,
          color = rnk_tier_lab,
          group = ""
        ),
        size = 3,
        alpha = 0.75
      ) +
      geom_point(
        data = data_tier1,
        aes(
          x = lon,
          y = lat,
          color = rnk_tier_lab,
          group = ""
        ),
        size = 5,
        alpha = 1
      ) +
      scale_alpha_continuous(range = c(0.5, 1)) +
      # scale_radius(range = c(1, 6)) +
      # scale_size_area(max_size = 6) +
      scale_color_manual(values = c("grey", "blue", "cyan")) +
      theme(legend.position = "bottom")

    if (add_labels) {
      ret <-
        ret +
        ggrepel::geom_label_repel(
          data = data_tier1,
          aes(
            x = lon,
            y = lat,
            label = school,
            group = ""
          )
        )
      # ggrepel::geom_label_repel(
      #   data = data_tier1,
      #   aes_string(
      #     x = "lon",
      #     y = "lat",
      #     label = col_label,
      #     group = ""
      #   )
      # )

    }

    ret
  }
```

```{r viz_map_tier3, echo = FALSE, include = TRUE, fig.show = "asis"}
viz_map_tier3 <-
  schools_stats %>%
  inner_join(
    schools_nces_join %>%
      distinct(school, county, lat, lon),
    by = c("school")
  ) %>%
  visualize_map_tier3_at(
    col = "prnk_sum",
    add_labels = TRUE,
    col_label = "school"
  ) +
  theme(
    panel.background = element_blank()
  ) +
  labs(
    title = "Most Dominant Schools",
    caption =
      str_wrap(
        paste0(
          "Dominance is evaluated according to total percent rank."
        ), 80
      )
  )
viz_map_tier3
```

We saw elsewhere that there is no
significant temporal trend for competititon types or competition level, but is
there one for school?


