---
author: ""
date: ""
title: ""
output:
  html_document:
    toc: false
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  cache = TRUE,
  # include = FALSE,
  # results = "markdown",
  # results = "hide",
  fig.align = "center",
  # fig.show = "asis",
  fig.width = 7,
  fig.height = 7,
  # out.width = 7,
  # out.height = 7,
  warning = FALSE,
  message = FALSE
)

```

```{r load, include = FALSE}
source(file.path("R", "config.R"))
schools_uil <-
  config$path_schools_uil %>%
  teproj::import_path_cleanly()

persons_uil <-
  config$path_persons_uil %>%
  teproj::import_path_cleanly()
```

```{r test}
persons_n4 <-
  persons_uil %>% 
  count(name, school, city, conf, comp, complvl, complvl_num) %>% 
  # filter(complvl == "State") %>% 
  filter(n == 4)

persons_n4_viz <-
  persons_uil %>% 
  semi_join(persons_n4) %>% 
  # filter(complvl == "State") %>% 
  unite(id, name, school, city, conf, comp, complvl, complvl_num, sep = "-", remove = FALSE) %>% 
  group_by(id, complvl) %>% 
  mutate(year_order = row_number(year)) %>% 
  ungroup()

viz_persons_n4 <-
  persons_n4_viz %>% 
  ggplot(aes(x = year_order, y = prnk)) +
  geom_line(aes(group = id), alpha = 0.1) +
  geom_smooth(method = "lm", color = "blue") +
  # geom_smooth(method = "loess", color = "red") +
  geom_point(
    data = 
      persons_n4_viz %>%
      group_by(complvl, year_order) %>%
      summarise(prnk_mean = mean(prnk)) %>% 
      ungroup(),
    aes(y = prnk_mean), size = 3, color = "blue") +
  facet_wrap(~complvl, scales = "free") +
  teplot::theme_te_facet()
viz_persons_n4
persons_multi <-
  persons_uil %>% 
  # filter(complvl == "District") %>% 
  count(name, school, city, conf, comp, complvl, complvl_num) %>% 
  filter(n >= 2)

persons_improve <-
  persons_uil %>% 
  semi_join(persons_multi) %>% 
  group_by(name, school, city, conf, comp, complvl, complvl_num) %>% 
  arrange(year, .by_group = TRUE) %>% 
  # mutate(order_byperson = row_number(desc(year))) %>% 
  # mutate(prnk_lag = lag(prnk)) %>% 
  mutate(improve = ifelse(prnk > lag(prnk), TRUE, FALSE)) %>% 
  ungroup() %>% 
  filter(!is.na(improve))
persons_improve_n <-
  persons_improve %>% 
  count(improve)
persons_improve_n_c <- tibble::deframe(persons_improve_n)
btest <-
  binom.test(persons_improve_n_c["TRUE"], sum(persons_improve_n_c)) %>% 
  broom::tidy()
btest
```


There's a lot to analyze with the Texas high school academic UIL data set.
Maybe I find it more interesting than otherse due to my personal experiences with
these competitions.

Now, after examining some of the biggest topics associated
with this data--including competitions, invidividuals, and schools--in a broad manner,
there are
some other things that don't necessarily fall into these categories that 
I think are worth investigating.

## Siblings


Let's look at the performance of siblings. [^siblings] Maybe this topic only came to mind for
me because I have brothers, on of who is my twin, but I think anyone can find
something interesting on this matter.

### Sibling Participation

[^sibling]:
The criteria for identifying siblings is same last name, school, and city.

So, let's start with something easy--which siblings competed together the most?

<i>
As in the analysis of individuals and schools, I'll include myself and my school in the tables and figures
where applicable.
</i>

```{r siblings_funcs, include = FALSE}
add_name_first_pair_col <-
  function(data = NULL) {
    data %>% 
      mutate(name_first_pair = paste0(name_first, " & ", name_first_sibling))
  }

get_siblings_by_at <-
  function(data = NULL,
           min = 2,
           max = Inf,
           cols_grp = NULL,
           prettify = TRUE) {

    data_proc <-
      data %>%
      group_by(!!!syms(cols_grp)) %>%
      mutate(rnk_max = rank(name_last, ties.method = "max")) %>%
      ungroup()

    data_filt <-
      data_proc %>%
      filter(rnk_max >= min, rnk_max <= max)

    ret <-
      data_filt %>%
      inner_join(data_filt,
                by = cols_grp,
                suffix = c("", "_sibling")) %>%
      filter(name_first != name_first_sibling)
    
    # ret <-
    #   ret %>% 
    #   distinct(!!!syms(cols_grp), .keep_all = TRUE)
    
    if(prettify) {
      ret <-
        ret %>% 
        add_name_first_pair_col() %>% 
        select(
          name_first_pair,
          name,
          name_first,
          name_last,
          name_first_sibling,
          everything()) %>%
        arrange(name_last, name_first, school)
      
      cols_drop <-
        setdiff(c(names(data), "name_first_pair", "name_first_sibling"), names(ret))
      
      ret <-
        ret %>%
        select(-one_of(cols_drop))
    }
      ret
  }

```


```{r siblings, include = FALSE}
siblings <-
  persons_uil %>%
  get_siblings_by_at(
    cols_grp = c("name_last",
                 "school",
                 "city",
                 "year",
                 "conf",
                 "complvl",
                 "comp",
                 "complvl_num")
  )
# NOTE: Need to fix double counting here.
siblings_n <-
  siblings %>% 
  group_by(name_last, name_first_pair) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(n) %>% 
  distinct(name_last, n, .keep_all = TRUE) %>%
  tetidy::rank_arrange_at("n")

```

```{r html_siblings_n, echo = FALSE, include = TRUE}
html_siblings_n <-
  siblings_n %>%
  create_kable_filt_at(
    config = config,
    col_rgx = "name_last",
  )
html_siblings_n
```
I was dissapointed to find that my twin brother and I were not at
the very top of this list. Nonetheless, we are fairly near the top, so I can take
some satisfaction in that.

I should say a couple of things about the previous table.

+ I don't explicilty try to filter for twins only, but it's reasonable to believe
that many, if not most, are twins.

+ The table enforces the criteria that the siblings compete
in the exact same competition--that is, a competition occurring in a given year and 
being of a same competition type and same competition level (as well as
the same conference and competition area, if applicable). The numbers
are inflated when not enforcing the criteria that the siblings must have competed
in the same competition type and level (nor conference and competition area), and even more so
when throwing out the criteria for same year.
For example, see the following counts for this relaxed criteria for sibling
participation.

```{r siblings_lax, include = FALSE, eval = FALSE}
# NOTE: Probably should remove this.
siblings_lax <-
  persons_uil %>%
  get_siblings_by_at(
    cols_grp = c("name_last", "school", "city")
  )

siblings_lax_n <-
  siblings_lax %>%
  group_by(school, name_last, name_first_pair) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  distinct(school, name_last, n, .keep_all = TRUE) %>%
  tetidy::rank_arrange_at("n")
siblings_lax_n
```

```{r html_siblings_lax_n, echo = FALSE, include = TRUE}
html_siblings_lax_n <-
  siblings_lax_n %>%
  create_kable_filt_at(
    config = config,
    col_rgx = "name_last",
  )
html_siblings_lax_n
```

### Sibling Performance

Participation in competitions is one thing, but what about sibling performance?
Let's use the same metric used elsewhere for ranking performance--percent rank of
scores summed accross all records (`prnk_sum`)--and see which sibling pairs
show up among the top.

```{r siblings_prnk, include = FALSE}
siblings_prnk <-
  siblings %>%
  select(matches("^name|^prnk")) %>%
  gather(prak_sum_type, value, matches("^prnk")) %>%
  group_by(name_last, name_first_pair) %>%
  summarise(sum = sum(value)) %>%
  ungroup() %>%
  distinct(name_last, sum, .keep_all = TRUE) %>%
  tetidy::rank_arrange_at("sum")
siblings_prnk
```

```{r html_siblings_prnk, echo = FALSE, include = TRUE}
html_siblings_prnk <-
  siblings_prnk %>%
  create_kable_filt_at(
    config = config,
    col_rgx = "name_last",
  )
html_siblings_prnk
```

It looks like the pairs at the top of these rankings based on score are fairly similar to the list
of pairs competing most frequently. (This is not too suprising
given that my choice of metric of ranking is based on a summed value that "rewards"
volume of participation rather than per-competition performance.) Again, my twin brother and I
appear near the top.

## My High School

Even though I highlighted my high school in my examination of schools and looked
at individual scores elsewhere, I did not
look at other individuals that have gone to my school. Perhaps it is a bit egotistical,
but I am interested in knowing how I compare with others that have attended my school (either
before, with, or after me).

```{r persons_n_filt, include = FALSE}
persons_n_filt <-
  persons_uil %>%
  group_by(name, school, conf) %>%
  summarise(n = n()) %>%
  ungroup() %>% 
  filter(str_detect(school, config$rgx_school_filt)) %>%
  tetidy::rank_arrange_at("n")
persons_n_filt
```

```{r html_persons_n_filt, echo = FALSE, include = TRUE}
html_persons_n_filt <-
  persons_n_filt %>% 
  create_kable_filt_at(
    config = config,
    col_rgx = "name"
  )
html_persons_n_filt
```



Alas, although my twin brother and I did not rank among the very top of the siblings
by participation and performance, we do appear among the top 
when evaluating only people from my high school. In my opinion, the sample size isn't so small
that this achievement is trivial. 


